Using Visual Analytics to Explain\\[1ex] Black-Box Machine Learning

### 2018-04-02 tex/introduction/machine_learning.tex
decision tree

\subsubsection{Decision Trees}
\todo{show image}
Instead of computing predictions with a mathematical formula, Decision Trees \cite{mlbook} determine predictions by following a sequence of tests on the data.
Decision Trees are trees where each node represents a test $x_i > t_i$, with $x_i$ as the value of the feature $i$ and $t_i$ as learned threshold.
The prediction algorithm starts at the root of the tree.
Depending on the outcome of the test in the root node the algorithm continues with the corresponding sub-tree recursively, until a leaf node is reached.
The leaf nodes contain the probabilities for the outcomes of the prediction task and are returned as result by the algorithm.

@@

However, Decision Trees cannot model relationships between features very well and have to approximate them.
For example, increasing the height of a person also increases that person's weight.
This means that whether a person is overweight, not only depends on the weight of the person, but also how tall that person is.
A person weighing $80\si{kg}$ can be considered overweight with a height of $1.7\si{m}$ but would not be with a height of $1.8\si{m}$.
Since a Decision Tree cannot model this relationship ($x_\text{mass} / x_\text{height}^2 \geq 25$) directly it has to approximate it by breaking the values of $x_\text{mass}$ and $x_\text{height}$ into intervals and checking them separately.
Because of this, Decision Trees are likely to over-fit on the training data.
Over-fitting happens when a machine learning model does not generalize well because of a pattern in the data that is only present in the training data.

>>>>

\subsubsection{Decision Trees}
Instead of computing predictions with a mathematical formula, Decision Trees \cite{mlbook} determine predictions by following a sequence of tests on the data.
Decision Trees are trees where each node represents a test $x_i > t_i$, with $x_i$ as the value of the feature $i$ and $t_i$ as learned threshold.
The prediction algorithm starts at the root of the tree.
Depending on the outcome of the test in the root node the algorithm continues with the corresponding sub-tree recursively, until a leaf node is reached.
The leaf nodes contain the probabilities for the outcomes of the prediction task and are returned as result by the algorithm.

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{tex/decisiontree}
    \caption[Example Decision Tree]{A possible decision tree for the Diabetes prediction example\footnotemark. Note, that the tree has interval checks to model the relationship between ``weight'' in \si{kg} and ``height'' in \si{m}.}
    \label{figs:dctree}
\end{figure}
\footnotetext{The decision tree or its parameters are \emph{not} based on data and serve only illustratory purposes.}

@@

However, Decision Trees cannot model relationships between features very well and have to approximate them.
For example, increasing the height of a person also increases that person's weight.
This means that whether a person is overweight, not only depends on the weight of the person, but also how tall that person is.
A person weighing $80\si{kg}$ can be considered overweight with a height of $1.7\si{m}$ but would not be with a height of $1.8\si{m}$.
Since a Decision Tree cannot model this relationship ($x_\text{mass} / x_\text{height}^2 \geq 25$) directly it has to approximate it by breaking the values of $x_\text{mass}$ and $x_\text{height}$ into intervals and checking them separately (see Figure~\ref{figs:dctree}).
Because of this, Decision Trees are likely to over-fit on the training data.
Over-fitting happens when a machine learning model does not generalize well because of a pattern in the data that is only present in the training data.
